
---
title: Quarto Basics
---

::: {.unit-overview}
# Overview

In this unit, we will cover the basics of what a Quarto document is and how to create and render it. 
:::

<!-- end unit-overview div -->

::: {.unit-goals}
# Goals

- Understand the basic anatomy of a Quarto document.
- Create and render documents to different output formats.


::: 
<!-- end unit-goals div -->

<!--
::: {.unit-video}
# Video, slides and audio transcript

Below is the video for this unit. Here are [the slides shown in the video](), and here is the [audio transcript](), as well as a [text version of the transcript]().

<iframe class="video" src="https://www.youtube.com/embed/tiwJzvqFgmY" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

::: 
-->
<!-- end unit-video div -->

::: {.unit-reading}
# Reading


## Structure of a Quarto Document

The Quarto software operates on a file called a **Quarto document** (`.qmd`). A standard Quarto file (`.qmd`) is a text document that consists of:

1. **YAML Front Matter**: Metadata and document configuration.
2. **Markdown Body**: Narrative text written in Markdown.
3. **Executable Code Chunks**: Embedded code blocks executed during rendering. This is optional, you can have a Quarto document without executable code chunks.


### Example of a Minimal Quarto Document

Here is an example of a minimal Quarto document:

````{verbatim}
---
title: "Simple Quarto Example"
format: html
---

The information above between the `---` lines is called the YAML front matter. It contains metadata about the document, such as the title and output format.

# This is a heading

The text in Quarto documents is formatted using Markdown syntax.

# Here is some code

A defining feature of Quarto is the ability to include executable code chunks within the document. Here is a simple example demonstrating the integration of R code and narrative.

This R code chunk generates a summary of the built-in `mtcars` dataset:

```{r}
summary(mtcars)
```
When you render this Quarto document, the R code will be executed, and the resulting summary statistics will be included in the output document.

````

As in the previous unit, start a new Quarto document in Positron and replace the content with the example above. After entering this code into the editor and saving it as `simple.qmd`, you can render it by typing `quarto render simple.qmd` into the terminal prompt (or use the `Preview` button.) An HTML file displaying your results will be created.


## Markdown Syntax

Quarto uses Markdown for formatting text. Markdown is what is called a markup language that allows you to format text using simple symbols. For example: 
- `#` for headings
- `**bold text**` for bold text
- `*italic text*` for italic text
- Lists using `-` or `*` for unordered lists and numbers for ordered lists


## Code Chunks

You can create Quarto documents that do not contain any code. However, a nice feature of Quarto is that it allows the combination of text and code, with code being executed during rendering and results being inserted into the final document. Code chunks are defined using triple backticks followed by the language identifier (e.g., `{r}` for R code): 

Add this code chunk to the Quarto document above 

````{verbatim}
```{r}
# This is an R code chunk
x <- seq(1, 10)
y <- x^2
plot(x, y, type = "b", main = "Plot of y = x^2", xlab = "x", ylab = "y")
```
````


Once you render it, you should see the code and the resulting plot in the output document, like this:

```{r}
# This is an R code chunk
x <- seq(1, 10)
y <- x^2
plot(x, y, type = "b", main = "Plot of y = x^2", xlab = "x", ylab = "y")
```



## YAML Front Matter

The YAML front matter is a section at the very top of a Quarto document enclosed by `---` lines. It contains metadata about the document and specifies how the document should be rendered. At a minimum, you should specify a title and output format. You can add all kinds of additional information to a YAML. That information controls various aspects of the document rendering and appearance.

### Output Formats

The starting point or input for Quarto is always one or multiple Quarto documents (`.qmd` files). From these source files, Quarto can generate output in various formats, including HTML, PDF, Word, and more.

For instance, if you want to create a Word document instead of HTML output for the example above, you would change the YAML from `format: html` to `format: docx`. Try doing that and re-rendering the document. You should now get a Word document as output instead of an HTML file. You should be able to find the word document in the same folder where your Quarto document is located. If fully configured, Positron might also give you the option to open the Word document.

Similarly, if you want to create a PDF document, you would change the YAML to `format: pdf`. Note that creating PDF documents requires a LaTeX installation on your computer. If you don't have LaTeX installed, Quarto will give you an error when you try to render to PDF. If you installed `tinytex` as explained in a previous unit, you should be able to produce a pdf file as output. Give it a try.



::: 
<!-- end unit-reading div -->

::: {.unit-summary}
# Summary

This unit introduced the basics of Quarto documents, including their structure, Markdown syntax, code chunks, and YAML front matter. You learned how to create and render Quarto documents in different output formats such as HTML, Word, and PDF.

::: 
<!-- end unit-summary div -->

::: {.unit-resources}
# Further Resources
* The [Quarto Website](https://quarto.org) provides comprehensive information on Quarto features and usage.
* The [Get Started section](https://quarto.org/docs/get-started/) of the Quarto website has a bit more information to get started, including a bit of 'behind the scenes' details.
* For documentation on Markdown syntax, check the [Markdown Basic site](https://quarto.org/docs/authoring/markdown-basics.html) of the Quarto website, and the references linked there.
* For more details on Code chunks, check the `Computation` section of the [Quarto Guide](https://quarto.org/docs/guide/).
* While there's no single place on the Quarto website that explains all of YAML, the `Documents` section of the [Quarto Guide](https://quarto.org/docs/guide/) is a good starting point.


::: 
<!-- end unit-resources div -->

::: {.unit-quiz}
# Test yourself

```{r quarto-basics-quiz}
#| echo: false
#| results: "asis"
quizzes <- list("quarto-basics-quiz-1.Rmd", "quarto-basics-quiz-2.Rmd", "quarto-basics-quiz-3.Rmd")
exams2forms::exams2forms(file = quizzes, title = 'quarto-basics-quiz')
```
::: 
<!-- end unit-quiz div -->

::: {.unit-exercise}
# Practice


::: 
<!-- end unit-exercise div -->
