
---
title: Basic Use
---

::: {.unit-overview}
# Overview

In this unit, you will learn the fundamental skills needed to create, edit, and render basic Quarto documents. You will explore the essential structure of a Quarto file and common workflows for generating outputs in different formats.
:::

<!-- end unit-overview div -->

::: {.unit-goals}
# Goals

By the end of this unit, you will:

- Understand the basic anatomy of a Quarto document.
- Create and render documents to multiple output formats.
- Utilize parameters and caching to improve efficiency and reproducibility.
::: 
<!-- end unit-goals div -->

<!--
::: {.unit-video}
# Video, slides and audio transcript

Below is the video for this unit. Here are [the slides shown in the video](), and here is the [audio transcript](), as well as a [text version of the transcript]().

<iframe class="video" src="https://www.youtube.com/embed/tiwJzvqFgmY" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

::: 
-->
<!-- end unit-video div -->

::: {.unit-reading}
# Reading


## Structure of a Quarto Document

The Quarto software operates on a file called a **Quarto document** (`.qmd`). A standard Quarto file (`.qmd`) is a text document that consists of:

1. **YAML Front Matter**: Metadata and document configuration.
2. **Markdown Body**: Narrative text written in Markdown.
3. **Executable Code Chunks**: Embedded code blocks executed during rendering. This is optional, you can have a Quarto document without executable code chunks.


### Example of a Minimal Quarto Document

Here is an example of a minimal Quarto document:

````markdown
---
title: "Simple Quarto Example"
format: html
---

Here is a simple example demonstrating the integration of R code and narrative.

```{r}
summary(mtcars)
```
````



After rendering (`quarto render simple.qmd`), an HTML file displaying your results will be created.



## Anatomy of a Quarto Document

A Quarto document typically consists of three main components:

1. **YAML Front Matter:** Metadata and output configurations defined at the beginning of the file.
2. **Markdown Text:** Narrative content formatted with Markdown syntax.
3. **Code Chunks:** Blocks of executable code embedded within the narrative.

Example:

````markdown
---
title: "Example Quarto Document"
format: html
---

This document integrates Markdown text with executable code.

```{r}
head(iris)
```
````

Render this document using: `quarto render example.qmd`

## Rendering Multiple Formats

Quarto allows the generation of multiple output formats from a single source file. You can specify multiple outputs in your YAML front matter:

```yaml
---
title: "Multi-format Document"
format:
  html: default
  pdf: default
---
```


To override and render a specific format:

`
quarto render multi-format.qmd --to pdf
`



## Caching and Freezing

Improve efficiency by caching results of time-consuming computations:

```yaml
execute:
  cache: true
```

Freeze documents to avoid unnecessary re-executions:


## Publishing Documents

To quickly publish your Quarto document as a website using GitHub Pages:



::: 
<!-- end unit-reading div -->

::: {.unit-summary}
# Summary

In this unit, you learned how to:

- Create a basic Quarto document with YAML front matter, Markdown content, and executable code.
- Render documents to multiple output formats from a single source file.
- Use parameters and caching to enhance efficiency and reproducibility.
- Publish your Quarto documents easily online.
::: 
<!-- end unit-summary div -->

::: {.unit-resources}
# Further Resources

- [Quarto Guide to Document Basics](https://quarto.org/docs/authoring/)
- [Parameters and Computations](https://quarto.org/docs/computations/parameters.html)
- [Caching and Freezing Execution](https://quarto.org/docs/computations/execution-options.html#caching)
- [Publishing with Quarto](https://quarto.org/docs/publishing/)
::: 
<!-- end unit-resources div -->

::: {.unit-quiz}
# Test yourself

```{r qsp-quiz}
#| echo: false
#| results: "asis"
exams2forms::exams2forms(file = "../../../assets/quizzes/dummy-quiz.Rmd", title = 'quiz-title', browse = FALSE)
```
::: 
<!-- end unit-quiz div -->

::: {.unit-exercise}
# Practice

### Practice Exercise

Create and render a Quarto document (`basic_use.qmd`) that outputs both HTML and PDF formats. Confirm both files are generated successfully.

### Substantial Exercise

Write a more complex Quarto document (`parametric_report.qmd`) that includes:

- YAML front matter specifying parameters.
- Markdown narrative introducing your analysis.
- Code chunks utilizing parameters to load data, perform analysis, and generate plots.
- Cached computations to optimize rendering.

Render the document with two different parameter sets and submit both sets of resulting HTML files, along with your source `.qmd` file.

::: 
<!-- end unit-exercise div -->
