---
title: Quarto and Positron
---

::: {.unit-overview}
# Overview

This unit explains how to use Quarto within the Positron IDE. 
:::

<!-- end unit-overview div -->

::: {.unit-goals}
# Goals

- Be able to create and render basic Quarto documents through Positron. 

::: 
<!-- end unit-goals div -->

<!--
::: {.unit-video}
# Video, slides and audio transcript

Below is the video for this unit. Here are [the slides shown in the video](), and here is the [audio transcript](), as well as a [text version of the transcript]().

<iframe class="video" src="https://www.youtube.com/embed/tiwJzvqFgmY" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

::: 
-->
<!-- end unit-video div -->

::: {.unit-reading}
# Reading

## Quarto and Positron  

Quarto documents are basically text documents. They have the file extension `.qmd` (Quarto Markdown). You can create and edit these files in any text editor, including Positron. 

## Creating Quarto documents with Positron

Start Positron. Then start a project (or open an existing one) as you learned in the _Brief Intro to Positron_ course. Next, create a new file by going to `File` -> `New File` -> `Quarto Document`. A new tab opens with a basic Quarto document template.

Delete the bit of text that's there, and replace it with this text.


```markdown
---
title: "Simple Quarto Example"
format: html
---

# This is a heading 

## This is a subheading

Here is text. **This is bold text.** 
Formatting follows the Markdown syntax.

## A code example

Here is a simple example demonstrating the integration of R code within the document.

```{r}
x <- seq(1, 10)
y <- x^2
plot(x,y)
```
The code is executed when the Quarto document is processed (rendered), and the resulting figure is automatically included in the output document.

```

Next, save the file in your project folder. You can name it anything, e.g. `simple-example.qmd`. The important part is to give it the `.qmd` ending to indicate that it's a Quarto file.

Positron supports editing Quarto files in two ways. One is called `Source`, which is where you see the actual text with Markdown formatting. The other one is called `Visual`, which gives you a preview of how the document will look when rendered. You can switch between these two modes by clicking the buttons in the top right corner of the editor window.

If you do switch back and forth, you might notice that sometimes the formatting of the source changes slightly. This should not matter and you should always be able to switch back and forth between source and visual views.


## Rendering

Once you have written your document and want to see the result, you can render it. The easiest way to do this is to click the `Preview` button at the top left of the editor window. Positron will execute the `quarto preview` command in the terminal (go there to see it). This both renders the document and opens the resulting file either inside the viewer in Positron or in your default software (e.g. a web browser for HTML output or MS Word for Word output).

Try this now. Click the `Preview` button. You should see the rendered document appear either in the viewer pane or in your web browser.

While it is convenient to press the `Preview` button in Positron, it's good to know and be able to run Quarto directly. Let's say you don't want the preview and instead just render the file, i.e. create the output file (here HTML) without opening it. 

You can do this by running Quarto through the terminal. Go to the terminal. It might still be busy from showing the preview. If you see something like "Listening on ..."  press `Ctrl + C` to stop it. You should then see the prompt, which ends with the `>` symbol. Now type the following command: 

```bash
quarto render name-of-file.qmd
```

As long as you are in the right folder, the one where you saved the Quarto file to, this will create the output file (here HTML) in the same folder. You should see the file show up in the 'Explorer' view in Positron (the top icon on the right sidebar). 


## Customizing Quarto Settings in Positron

Positron comes with the Quarto extension pre-installed, providing seamless integration for Quarto document creation and rendering. You can change settings for Quarto as you learned previously in the Positron Extensions unit of the _Brief Intro to Positron_ course.



## Debugging and Troubleshooting

If you encounter issues when using Quarto, you will usually see an error message in the `Terminal` pane that explains what might be going wrong. Quite often, rendering fails if the front-end YAML is not properly formatted.



::: 
<!-- end unit-reading div -->

::: {.unit-summary}
# Summary

Quarto is nicely integrated into Positron, allowing you to create, edit, and render Quarto documents directly within the IDE.

::: 
<!-- end unit-summary div -->

::: {.unit-resources}
# Further Resources

::: 
<!-- end unit-resources div -->

::: {.unit-quiz}
# Test yourself

```{r qsp-quiz}
#| echo: false
#| results: "asis"
#exams2forms::exams2forms(file = "../../../assets/quizzes/dummy-quiz.Rmd", title = 'quiz-title', browse = FALSE)
```
::: 
<!-- end unit-quiz div -->

::: {.unit-exercise}
# Practice

* Find a Quarto file somewhere online. Copy to your computer, open it in Positron, and render it. 
* Open the Quarto settings and change some of the options. Then create and render a new Quarto document and see if or how the changed settings affect the new document. 

::: 
<!-- end unit-exercise div -->
