---
title: Basic Positron Use 
---

::: {.unit-overview}
# Overview

In this unit, we'll cover some of the basics of using Positron. 

:::
<!-- end unit-overview div -->

::: {.unit-goals}
# Goals
* Be familiar with basic ways to use Positron.

:::
<!-- end unit-goals div -->

<!--
::: {{.unit-video}}
# Video, slides and audio transcript

Below is the video for this unit. Here are [the slides shown in the video](), as well as a [text version of the transcript]().

<iframe class="video" src="https://www.youtube.com/embed/VIDEO_ID" title="Video placeholder" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

:::
-->
<!-- end unit-video div -->

::: {{.unit-reading}}
# Reading

## Introduction

At the heart of every IDE, such as Positron, is the editor. Positron allows you to edit files efficiently, and if it is code, you can also run it. We'll explore some basic examples in this short unit.


## Editing a text file

This is the most boring use of the editor. Just create a new file, or open an existing text file and edit it.
A text file doesn't necessarily need to be plain text. You can for instance open and edit a CSV file as well. 

Depending on the type of file you are editing, Positron will try to provide some features that make it easier to work with it. For instance, if you open an R script, it will highlight the syntax and provide some code completion features. Syntax highlighting exists for many languages, not just R. As you work with the editor and learn more features, you will continue to find features that make your workflow easier and more efficient.  


## Editing and running code

Positron is meant for working with code. It is especically designed to work well with R and Python, but other languages are also supported, many through extensions. Let's explore a small example with R. Go to "New" -> "New File" and pick "R Script". An empty file will open in the editor. Write a few lines of R code, for instance the following

```{r}
#| echo: true
#| eval: false
x = seq(1,100)
y = x^2
plot(x,y)
```

Next, save the file somewhere. If Positron was 'just an editor', we would now have to switch over to R and run the code there. But with an IDE, we can do it all in one place. You should see a small triangle symbol in the top right corner of the editor. Press it to run the code. This will open the R console in the bottom window, execute the 'source' command to run the code, and show the resulting plot in the bottom window. In the top right window, you can also see the variables that were generated.

Let's say you want to have a line instead of circles for the plot but don't remember the command. In the R console, you can type `?plot` to call the help file for the `plot` function, which will show in a separate pane.

Once you found what you are looking for, change your code so it says `plot(x,y,type='l')` and run again. This time, instead of pressing the 'run' button, try to execute the code one line at a time. You can do that by starting in the first line and pressing `Ctrl/Cmd + Enter`. The first line will be executed, and you can then continue to the next line by pressing `Ctrl/Cmd + Enter` again. You will see the new plot show up. You should also see that the old plot is still there, you can go back to it with the arrows above the plot. The menu above the plot also allows you to save any of them, and do other things. It is generally not advised to do any steps manually, you should instead have code do all of it to ensure reproducibility. But in situations where that is not too important, these little helpers are very useful.

<!-- You can of course also do the same for other programming languages. We'll look at an example for Julia in the [Introduction to Julia course](/courses/course-landing-pages/fc5-intro-julia.qmd) -->


## Working with Quarto documents

Let's try a brief quarto document. Quarto itself is discussed in our [Introduction to Quarto course](/courses/course-landing-pages/fc2-intro-quarto.qmd), so we won't talk about it much here, just showcase how to work with it in Positron. 

Start another new file, and this time from the menu, pick `Quarto Document` towards the bottom. Another editor window should open with an almost empty area. There will be a few lines at the top indicating a Quarto document.

Give it a title, write a heading with `# Heading`, then add some text.
Next take the code from the previous example and insert it into the Quarto file as code chunk, like this

````{verbatim}
```{r}
#| echo: true
#| eval: true
x = seq(1,100)
y = x^2
plot(x,y)
```
````

Then add a bit more text after this code chunk.
Finally, safe the file as `.qmd` file.

Then press the 'preview' button above the editor window. Quarto should run and compile/render the document. If things work right, you should see a preview of your Quarto document show up in the viewer pane or open in an external browser, depending on your settings for the Quarto extension.


:::
<!-- end unit-reading div -->

::: {.unit-summary}
# Summary

You saw two examples illustrating the basics of working with Positron. There is much more to discover. We'll cover some of it in other courses and there's much more that you might want to learn on your own. Fortunately, Positron is rather user-friendly, and since it is based on VS Code, which has been around for a long time and is very popular, there is a lot of documentation on it. 

:::
<!-- end unit-summary div -->

::: {.unit-resources}
# Further Resources


:::
<!-- end unit-resources div -->

::: {{.unit-quiz}}
# Test yourself

```{{r qsp-quiz}}
#| echo: false
#| results: "asis"
exams2forms::exams2forms(file = "../../../assets/quizzes/dummy-quiz.Rmd", title = 'quiz-title', browse = FALSE)
```
:::
<!-- end unit-quiz div -->

::: {{.unit-exercise}}
# Practice

* 


:::
<!-- end unit-exercise div -->
